---
title: "Merger Arbitrage"
author: "LB"
date: "10/7/2020"
output: html_document
---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)
```


```{r load-libraries, include=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(ggthemes)
library(lubridate)
library(here)
library(skimr)
library(janitor)
library(httr)
library(readxl)
library(vroom)
library(data.table)
library(infer)
library(ggridges)
library(viridis)
library(tidyquant)
library(rvest)    # scrape websites
library(purrr)  
library(lubridate) #to handle dates
library(ggrepel) # to avoid loverlapping legends
library(ggplot2)
library(broom)
library(huxtable)
```

## Benchmark data

Risk-free Rates

```{r}
tbill <- vroom("https://fred.stlouisfed.org/graph/fredgraph.csv?bgcolor=%23e1e9f0&chart_type=line&drp=0&fo=open%20sans&graph_bgcolor=%23ffffff&height=450&mode=fred&recession_bars=on&txtcolor=%23444444&ts=12&tts=12&width=1168&nt=0&thu=0&trc=0&show_legend=yes&show_axis_titles=yes&show_tooltip=yes&id=DTB3&scale=left&cosd=2009-08-01&coed=2020-10-09&line_color=%234572a7&link_values=false&line_style=solid&mark_type=none&mw=3&lw=2&ost=-99999&oet=99999&mma=0&fml=a&fq=Daily&fam=avg&fgst=lin&fgsnd=2020-02-01&line_index=1&transformation=lin&vintage_date=2020-10-14&revision_date=2020-10-14&nd=1954-01-04") %>%
  mutate(DTB3 = DTB3/100) %>%
  rename("date" = "DATE",
         "T_bill" = "DTB3") %>% 
  mutate(T_bill_d = (1 + T_bill)^(1/252)-1,
         date = as.Date(date,"%Y-%m-%d"))
```

Merger Deals

```{r, get tickers}

# we need a vector of strings with just the 30 tickers + SPY
stocks <- 
  data.frame(acquirer = c("ADS","AAPL"),
             target = c("NKE", "GOOGL"),
             acquirer_stocks = c(-1, -2),
             target_stocks = c(5, 6),
             announce_date =c(
               "2011-08-06",
               "2011-09-07"),
             close_date =c(
               "2011-09-08",
               "2012-09-05"))

```

## Get Stock Data and build Dataframes

```{r get_price_data, message=FALSE, warning=FALSE, cache=TRUE}
# Notice the cache=TRUE argument in the chunk options. Because getting data is time consuming, # cache=TRUE means that once it downloads data, the chunk will not run again next time you knit your Rmd

stock_data_acquirer <- stocks %>% 
  select(acquirer) %>% 
  tq_get(get  = "stock.prices",
         from = "2009-01-01",
         to   = "2020-08-31") %>%
  group_by(acquirer) %>% 
  select(acquirer, date, close) %>% 
  mutate(close_acquirer = sprintf("%.2f",close))

stock_data_acquirer_target <- left_join(stock_data_acquirer,stocks , by=c("acquirer"))


stock_data_returns_acquirer <- stocks %>% 
  select(acquirer) %>% 
  tq_get(get  = "stock.prices",
         from = "2009-01-01",
         to   = "2020-08-31") %>%
  group_by(acquirer) %>% 
  select(acquirer, date, close) %>% 
  group_by(acquirer) %>%
    tq_transmute(select     = close, 
                 mutate_fun = periodReturn, 
                 period     = "daily", 
                 col_rename = "R_d_acquirer")

df_acquirer <- left_join(stock_data_acquirer_target,stock_data_returns_acquirer , by=c("acquirer","date"))

stock_data_target <- stocks %>% 
  select(target) %>% 
  tq_get(get  = "stock.prices",
         from = "2009-01-01",
         to   = "2020-08-31") %>%
  group_by(target) %>% 
  select(target, date, close) %>% 
  mutate(close_target = sprintf("%.2f",close))

stock_data_returns_target <- stocks %>% 
  select(target) %>% 
  tq_get(get  = "stock.prices",
         from = "2009-01-01",
         to   = "2020-08-31") %>%
  group_by(target) %>% 
  select(target, date, close) %>% 
  group_by(target) %>%
    tq_transmute(select     = close, 
                 mutate_fun = periodReturn, 
                 period     = "daily", 
                 col_rename = "R_d_target")

df_target <- left_join(stock_data_target,stock_data_returns_target , by=c("target","date"))

df_returns <- left_join(df_acquirer,df_target , by=c("date","target")) 

df_returns <- df_returns %>% 
    mutate(announce_date = as.Date(announce_date,"%Y-%m-%d"),
           close_date = as.Date(close_date,"%Y-%m-%d"),
           standard_date = date - announce_date) %>% 
  group_by(acquirer) %>% 
  filter(standard_date >= -2,
         date <= close_date)

df_returns <- df_returns %>% 
  mutate(close_acquirer = parse_number(close_acquirer),
         close_target = parse_number(close_target)) %>% 
  group_by(acquirer) %>% 
  mutate(period = row_number(),
         period = period + as.numeric(time_length( min(standard_date), "days")) -1,
         R_d_combined = R_d_target - R_d_acquirer) %>% 
  select(-announce_date, -close_date,
         -standard_date,
         -acquirer_stocks, -target_stocks)

df_returns <- df_returns[,c("date",
                            "period",
                            "acquirer",
                            "target",
                            "close_acquirer",
                            "close_target",
                            "R_d_acquirer",
                            "R_d_target",
                            "R_d_combined")]


df_standard_returns <- df_returns %>% 
  group_by(period) %>% 
  # Get average returns across deals
  summarise(return_acquirer = mean(R_d_acquirer),
            return_target = mean(-R_d_target),
            return_combined = mean(R_d_combined),
            # set up cumulative return field
            cum_return_acquirer = 1,
            cum_return_target = 1,
            cum_return_combined = 1) %>% 
  # calculate cumulative return per day
  mutate(cum_return_acquirer = ifelse(is.na(lag(cum_return_acquirer)),1,lag(cum_return_acquirer)) * (1+return_acquirer),
         cum_return_target = ifelse(is.na(lag(cum_return_target)),1,lag(cum_return_target)) * (1+return_target),
         cum_return_combined = ifelse(is.na(lag(cum_return_combined)),1,lag(cum_return_combined)) * (1+return_combined))

# Create Time-series 
df_ts_returns <- left_join(tbill, df_returns, by ="date") %>% 
  # calculate average returns for all strategies by date
  group_by(date) %>% 
  summarise(T_bill_d = mean(T_bill_d),
            R_d_acquirer = mean(ifelse(period >= 0, R_d_acquirer,NA)),
            R_d_target = mean(ifelse(period >= 0, R_d_target,NA)),
            R_d_combined = mean(ifelse(period >= 0, R_d_combined,NA))) %>% 
  # create merger arbitrage strategy which moves into risk free bonds if no current deal
  mutate(R_d_strategy = ifelse(is.na(R_d_combined), T_bill_d, R_d_combined))




# reformat geometric formula to exponential formula https://en.wikipedia.org/wiki/Geometric_mean#:~:text=2%2C%20that%20is%2C-,.,a%20financial%20investment%20over%20time.
gm_mean = function(x, na.rm=TRUE){
  exp(sum(log(1+x)) / length(x)) -1}

```



# Analysis


Index

```{r}
benchmark <- 
  data.frame(index = c("MNA", "^RUA", "^FTSE","^FTAS", "^IXIC","NYA", "^DJI", "^GSPC"),
             description = c("IQ Merger Arbitrage ETF", "Russell_3000", "FTSE 100", "FTSE All Share","NASDAQ_Composite", "NYSE Composite", "Dow Jones Industrial Average", "SP_500")) %>% 
  tq_get(get  = "stock.prices",
         from = "2009-01-01",
         to   = "2020-08-31") %>%
  select(index, description, date, close) %>%
  mutate(index_return = close/lag(close) - 1)
```

## CAPM
```{r}
CAPM <- df_ts_returns %>%
  left_join(benchmark, by = "date") %>%
  select(date, R_d_acquirer, R_d_target,R_d_combined, R_d_strategy,T_bill_d, description, index_return) %>%
  filter(description == "Russell_3000") %>%
  pivot_wider(values_from = index_return, names_from = description) %>%
  mutate(Rm_Rf = Russell_3000 - T_bill_d,
         Rs_Rf = R_d_strategy - T_bill_d)

ggplot(CAPM, aes(x = Rm_Rf, y = Rs_Rf)) +
  geom_point() 

CAPM_regression <- lm(Rs_Rf ~ Rm_Rf, data = CAPM)

CAPM_regression

huxreg(CAPM_regression,
       statistics = c('#observations' = 'nobs', 
                      'R squared' = 'r.squared', 
                      'Adj. R Squared' = 'adj.r.squared', 
                      'Residual SE' = 'sigma'), 
       bold_signif = 0.05, 
       stars = NULL
)


```
# To Dos
## Get standard dates into proper series. - Dung
## Add portfolio set-up and portfolio close price - Leif
## calculate portfolio daily returns  -> mutate( ret = close_price/lag(close_price) - 1) - Leif
## calculate daily returns across deals with gm_function - Leif
## plot results - Dung
## measure performance. Regression with Market index -> calculate Alpha and Beta - Dung
